---
title: "Custom distance measures"
author: Maciej Eder
date: 27/08/2015
output: 
  github_document:
  #  highlight: pygments
    fig_width: 7
    fig_height: 5
bibliography: bibliography.bib
csl: digital_humanities_abstracts.csl
---



```{r setup, include = FALSE}
# set global chunk options

library(stylo)
#library(knitr)

# using a cache folder
# opts_chunk$set(cache = TRUE)
```



Distance measures in stylometry... This is definitely the topic one never gets bored with. I’ve been promising myself to write a longer piece on this one day: not this time, though. In this short post, I’m going to introduce a functionality of the R package `stylo` (ver. >= 0.6.0) that allows for testing _any_ distance.

Apart from the already-implemented distance measures (most of them can be conveniently selected via GUI), the package `stylo` features a socket for plugging in your own custom distances. In short, a bit of coding combined with some expertise in maths – that’s basically all we need. We have to design a function which takes a table with frequencies as an input parameter, and returns a square table of distances: it can be either an object of the generic class `dist` (the lower triangle of the distance matrix stored by columns in a vector), or a regular full matrix,  symmetric across the diagonal. Whatever. 

Suppose you want to test the distance discussed in a great paper by Jannidis, Schoch, Pielstrom & Vitt [@jannidis_improving_2015] presented at DH2015. This is a regular Cosine Distance applied to z-scored data. Interestingly enough, this measure has been (sort of) introduced in an earlier study [@smith_improving_2011], but never tested before the Würzburg guys took over. First, we prepare a custom function (this is a simple version: a real function should check if the input dataset can be further processed, if it is a matrix, etc.). Type the following code:


``` {r eval = FALSE}
my.cosine.distance = function(x){
    
    # z-scoring the input matrix of frequencies
    x = scale(x)
    
    # computing cosine dissimilarity
    y = as.dist( x %*% t(x) / (sqrt(rowSums(x^2) %*% t(rowSums(x^2)))) ) 
    
    # then, turning it into cosine similarity
    z = 1 - y
    
    # getting the results
    return(z)
}
```


Once the above code is copy-pasted to the R console, it becomes a new function named `my.cosine.distance()`, and it becomes visible for other R objects. The function, however, is not persistent: the paste-copying step has to be repeated every time a new R session launched. 

Having completed the above step, we're all set. Now, one can use the tailored distance function with any of the main functions of the package `stylo`. Note the following examples:


``` {r eval = FALSE}
stylo(distance.measure = "my.cosine.distance")

classify(distance.measure = "my.cosine.distance")

rolling.classify(distance.measure = "my.cosine.distance")
```

The Cosine Delta (aka Würzburg Delta) is but one example of replacing the original kernel of the Delta method with a custom distance. What about trying something else? Assume that you plan to test if the Entropy Distance outperforms other similarity measures. It has been reported in the literature [@juola_controlled-corpus_2005] that entropy-based distances are generally accurate in stylometry. Let’s define a tailored function:


``` {r eval = FALSE}
dist.entropy = function(x) {
    A = t(t(x + 1) / colSums(x + 1))
    B = t(t(log(x + 2)) / -(colSums(A * log(A))))
    y = dist(B, method="manhattan")
    return(y)
}
```

The next step is rather obvious, given the already-discussed examples:

``` {r eval = FALSE}
stylo(distance.measure = "dist.entropy")
```


Now, what about an inverse correlation distance? It has been successfully applied in a cross-language benchmark study distances in stylometry [@forsyth_document_2014], in which, by the way, a few other interesting measures have been tested. The following code was contributed by Richard Forsyth:


``` {r eval = FALSE}
##  Additional distance function for Stylo() :
##  submitted by R.S. Forsyth.
##  First version : 30/10/2013
##  Last revision : 30/10/2013

cordist = function (freqvec,cols=89,usemeth="spearman") {
	##  takes 2D vector of word frex [,1:cols] (table.with.all.freqs);
	##  returns a distance matrx 2b used in clustering or MDS.
	##  Distance index is inverse correlation (Rank.Corr default).

	if (cols < 2)  stop("Too few cols!!")
	##  table.with.all.freqs usually has too many cols;
	##  don't know where correct number to be used is saved;
	##  therefore currently needed as function argument.
	tranfrex = t(freqvec[,1:cols])  ##  columns are features (words/grams)
	dmat = 1-cor(tranfrex,meth=usemeth)
	##  meth cd also be "kendall" or "pearson" (not recommended).

	return  (as.dist(dmat))
	##  names seem to stay attached, as attr(*,"Labels").

}  ##  ready for hclust() or sammon() scaling.
```

Etc. etc. etc. It is claimed that over 5,000 distance measures have been introduced so far in exact sciences [@moisl_cluster_2014]. Try them all!




## References

